---
title: "Section 2: Motivation and Grammar of Graphics"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
## install.packages("palmerpenguins")
library(palmerpenguins)
library(tidyverse)
penguins

## single quantitative frequency plot
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly() 
## single categorical barplot
ggplot(data = penguins, aes(x = species)) +
  geom_bar()
## two quant scatterplot
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth()
## two cat heatmap
ggplot(data = penguins, aes(x = species, y = island)) +
  geom_bin_2d()
```

## Review Exercises

```{r}
## 1
ggplot(data = penguins, aes(x = species, y = bill_length_mm)) +
  geom_violin()

## 2
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly(aes(colour = species)) 

ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly(colour = "darkorange3") 

## 3
ggplot(data = penguins, aes(x = bill_length_mm,
                            y = bill_depth_mm,
                            colour = species)) +
  geom_point() +
  geom_smooth()

## 4

ggplot(data = penguins, aes(x = bill_length_mm,
                            y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  geom_smooth()

```

```{r}
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_histogram(colour = "black", fill = "white") +
  facet_wrap(~ species, ncol = 1)

penguins <- penguins %>% mutate(random_var = sample(c("A", "B", "C", "D", "E", "F", "G"),
       size = nrow(penguins),
       replace = TRUE))
ggplot(data =penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = random_var)) +
  geom_point() 

ggplot(data =penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~random_var)

ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ species)
```

```{r}

ggplot(data = penguins, aes(x = species)) +
  geom_bar(stat = "count") ## don't need stat = "count" because it's the default.

penguins_count <- penguins %>% group_by(species) %>%
  summarise(ncount = n())

ggplot(data = penguins_count, aes(x = species)) +
  geom_bar(stat = "count")



ggplot(data = penguins, aes(x = species, fill = sex)) +
  geom_bar(position = "stack") + ## don't need "stacked" because it's default
  scale_fill_viridis_d()

penguins_nomiss <-penguins %>% filter(!is.na(sex))

ggplot(data = penguins_nomiss, aes(x = species, fill = sex)) +
  geom_bar(position = "dodge") + ## don't need "stacked" because it's default
  scale_fill_viridis_d()

ggplot(data = penguins, aes(x = bill_length_mm,
                            y = bill_depth_mm)) +
  geom_point(stat = "identity", position = "identity") +
  facet_null() +
  coord_cartesian()

ggplot(data = penguins, aes(x = species)) +
  geom_bar() +
  coord_polar()
ggplot(data = penguins, aes(x = species)) + 
  geom_bar() +
  coord_flip()
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm,
                            colour = species)) +
  geom_point() +
  geom_smooth()
```

```{r}
library(tidyverse)
category <- rep(c("Group 1", "Group 2", "Group 3", "Group 4", "Group 5"), 7)
time <- rep(c(2000, 2001, 2002, 2003, 2004, 2005, 2006), each = 5) 
response <- rpois(length(category), 30)
df <- tibble(category, time, response)
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category)) +
  labs(y = "count") +
  scale_fill_viridis_d()

ggplot(data = df, aes(x = time, y = response)) +
  geom_line(aes(colour = category)) +
  scale_colour_viridis_d()
```

## Section 2.3

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm,
                            colour = species)) +
  geom_point() +
  geom_smooth() +
  theme(plot.background = element_rect(fill = "yellow"),
        axis.ticks.length=unit(1.1, "cm"),
        panel.grid = element_line(colour = "brown", size = 1.5),
        aspect.ratio = 5,
        legend.position = c(0, 0),
        legend.direction = "horizontal",
        legend.key = element_rect(fill = "darkorange3", colour = "blue")) +
  scale_colour_manual(values = c("yellow", "green", "purple"))
```






Exercise 1 Questions

Do different rules apply for animated/interactive graphs in a web app?

What are some of the first things you should look at on a graph when reading it to make sure that it is actually telling you what it is supposed to?

Is there ever a reasonable time to intentionally make bad graphs?

if Minard made his famous graphic by hand, or what kind of tools he used, since he did not have the technology we have today

Are there any situations in which adding dimensions to a visualization would be useful?

Is there actually a time and place for 3d graphics at all? Is it time to remove some portion of them from our data visualization arsenal altogether?

I would like to get the data-to-ink ratio explained a little more. I didn't quite grasp it

questions about 3d graphics:

https://shiny.rit.albany.edu/stat/mod1/

When would we use area graphs?



Should all graphs start at 0? scatterplots? bar charts? Line charts? Boxplots?

<https://flowingdata.com/2015/08/31/bar-chart-baselines-start-at-zero/> 
