---
title: "Section 3: Applied Concepts"
format: html
---


```{r}
library(tidyverse)
library(palmerpenguins)
theme_set(theme_minimal())

penguins_sum <- penguins |> group_by(species, year) |>
  summarise(n_penguins = n()) |>
  mutate(year = factor(year))

ggplot(data = penguins_sum, aes(x = year, fill = species)) +
  geom_col(aes(y = n_penguins)) +
  theme_minimal() +
  scale_fill_viridis_d()
```

The line plot is overall preferable because it's hard to see trends in the species of penguins that are not the Gentoo species (which is the baseline).

Example 2.

The first is preferable because, if our visualization encodes area or length, the y-axis should start at 0.

Example 3.

The first is preferable because including 0 as a y limit for this type of graph gives a lot of wasted space.

Example 4.

Pair 1

The first is preferable because it better utilizes the position of points so we can easily see which species has high vs. low bill depths, etc.

Pair 2. The second is preferable because its hard for humans to accurately perceive 10 different color shades.

```{r}
penguins <- penguins |> mutate(species_island_sex = interaction(species, 
                                                                island,
                                                                sex))
ggplot(data = penguins |>
         filter(!is.na(sex)), aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species_island_sex)) +
  scale_colour_viridis_d() +
  theme_minimal()
ggplot(data = penguins |> filter(!is.na(sex)), 
                                 aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap(~ species_island_sex) +
  theme_minimal()

ggplot(data = penguins |> filter(!is.na(sex)),
       aes(x = bill_length_mm,
           y = bill_depth_mm)) +
  geom_point(aes(colour = sex, shape = species)) +
  facet_wrap(~ island)

ggplot(data = penguins |> filter(!is.na(sex)),
       aes(x = bill_length_mm,
           y = bill_depth_mm)) +
  geom_point(aes(colour = sex)) +
  facet_grid(species ~ island)
```

Example 5.

The bar plot is stronger because it effectively uses position of the bars while the areas of the pie chart are harder to accurately interpret.

Example 6. ??? In the first, it's easier to distinguish the region while the second has a higher data to ink ratio.

## Exercises

Exercise 1. 

a. sequential (number of cases is clearly ordered: more cases would be a more "pop" colour)

b. diverging (0 change would be given a neutral colour)

c. qualitative

d. diverging (0 change would be a neutral colour)

Exercise 2.

```{r}
ggplot(data = penguins, aes(x = bill_length_mm,
          y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  scale_colour_viridis_d() ## d is for discrete variable

ggplot(data = penguins, aes(x = bill_length_mm,
          y = bill_depth_mm)) +
  geom_point(aes(colour = body_mass_g)) +
  scale_colour_viridis_c() ## c is for continuous

ggplot(data = penguins, aes(x = bill_length_mm,
          y = bill_depth_mm)) +
  geom_point(aes(colour = body_mass_g)) +
  scale_colour_viridis_b() ## b is binning a continuous
```

Exercise 3.

```{r}
hpi_df <- read_csv(here("data/hpi-tidy.csv"))
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing, colour = Region)) +
  geom_point() +
  theme(plot.background = element_rect(fill = "yellow"),
        axis.text = element_text(size = 4,
                                 family = "courier",
                                 colour = "green"),
        axis.line = element_line(arrow = arrow()),
        panel.grid = element_blank())
```
