---
title: "`tidyverse` Review"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aside: Baseline of Zero?

```{r}
library(palmerpenguins)
ggplot(data = penguins, aes(x = bill_length_mm,
                            y = bill_depth_mm)) +
  geom_point() +
  xlim(c(0, 60)) +
  ylim(c(0, 25))

```

## Section 5.1: Billboard

```{r}
## install.packages("billboard")
library(billboard)
head(wiki_hot_100s)
tail(wiki_hot_100s)

wiki_top15 <- wiki_hot_100s %>% 
  as_tibble() %>% 
  filter(year <= 2009 & year >= 2000) %>%
  group_by(artist) %>%
  summarise(nsongs = n()) %>%
  arrange(desc(nsongs)) %>%
  slice(1:15) %>%
  mutate(artist = fct_reorder(artist, nsongs))


## exercise 3
wiki_hot_100s %>% as_tibble() %>% 
  mutate(artist2 = str_remove(artist, 
                              pattern = " featuring .*")) %>%
  tail() %>%
  print(width = Inf)

ggplot(data = wiki_top15, aes(x = artist,
                              y = nsongs)) +
  geom_col() +
  coord_flip() 

## exercise 4
ggplot(data = wiki_top15, aes(x = artist,
                              y = nsongs)) +
  geom_point() +
  geom_segment(aes(x = artist, xend = artist,
                   y = 0, yend = nsongs)) +
  coord_flip()

max(wiki_hot_100s$year)
library(rvest)
library(tidyverse)

## provide the URL and name it something (in this case, url).
## paste0 pastes together the base URL and the year into a single string:
## this will be useful in a moment
year <- 2017
url <- paste0("https://en.wikipedia.org/wiki/Billboard_Year-End_Hot_100_singles_of_", year)

## convert the html code into something R can read
h <- read_html(url)

## grabs the tables
tab <- h %>% html_nodes("table")
df <- tab[[1]] %>% html_table() %>%
  mutate(year = 2017)
df
get_wiki_100 <- function(year) {
  
  ## same code as before, replacing 2017 with year.
  url <- paste0("https://en.wikipedia.org/wiki/Billboard_Year-End_Hot_100_singles_of_", year)
  
  h <- read_html(url)
  
  tab <- h %>% html_nodes("table")
  df <- tab[[1]] %>% html_table() %>%
    mutate(year = year)
  
  ## tell our function to return the dataframe `df`
  return(df) 
}
get_wiki_100(year = 2017)
library(purrr)
year_list <- list(2017, 2018, 2019, 2020, 2021)
year_list

df_all <- map(year_list, get_wiki_100)
df_all ## a list of data frames, one for each year
df_2017_present <- bind_rows(df_all)
df_2017_present <- df_2017_present %>%
  mutate(Title = str_remove_all(Title, pattern = "\"")) %>% ## get rid of \ in title
  rename(no = No., 
         title = Title, 
         artist = `Artist(s)`) ## make column names match with billboard package

wiki_tibble <- as_tibble(wiki_hot_100s) %>% ## convert billboard data to tibble
  mutate(year = as.numeric(year),
         no = as.integer(no)) ## change variable types to match with scraped data

hot100_df <- bind_rows(wiki_tibble, df_2017_present)
```


```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")
hpi_df
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()

hpi_us <- hpi_df %>%
  filter(Country == "United States of America" |
           Country == "Kenya" |
           Country == "Brazil")

hpi_us
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label(data = hpi_us, aes(label = Country)) ## specify
## data = hpi_us so geom_label only uses the observation in hpi_us
library(ggrepel)
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) ## create a second point that is an open circle (shape = 1) with a larger size (size = 3) to surround the United States point on the scatterplot
## install.packages("plotly")
library(plotly)
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()
ggplotly(plot1)
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing,
                                   label = Country)) +
  geom_point()
ggplotly(plot1, tooltip = "label")
```

## Exercise 3

Advantages: interesting, allows user to explore
Disadvantages: requires a computer, adds extra "stuff", bad for making a particular point about the graph (like Western countries have high footprint)
```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) +
  labs(title = "Countries with a Higher Ecological Footprint Tend to Have Citizens with Higher Wellbeing", ## add title
    subtitle = "Wellbeing is on a 1-10 scale", ## add subtitle (smaller text size than the title)
    caption = "Data Source: http://happyplanetindex.org/countries", ## add caption to the bottom of the figure
    x = "Ecological Footprint", ## change x axis label
    y = "Wellbeing") ## change y axis label
ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_brewer(palette = "Dark2")


ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_viridis_d(option = "plasma")
?scale_colour_viridis_d
```

## Exercise 6

`_d` is for discrete values, `_c` is for continuous variables, and `_b` is for binning a continuous variable.
```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears)) +
  geom_point() +
  facet_wrap( ~ Region)
library(palmerpenguins)
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) ## colour is good enough here
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap( ~ species) ## faceting probably unnecessary: colour is better
```

Faceting is more useful if:

* there are "more" categories and
* if there is a lot of overlap between groups

```{r}
colour_bad <- tibble(x = rnorm(500, 0, 1), y = rnorm(500, 0, 1),
       groupvar = c(rep("A", 50), rep("B", 50),
                    rep("C", 50), rep("D", 50),
                    rep("E", 50), rep("F", 50), rep("G", 50),
                    rep("H", 50), rep("I", 50), rep("J", 50)))

ggplot(data = colour_bad, aes(x = x, y = y, colour = groupvar)) +
  geom_point() + ## can't distinguish anything really: colour is bad
  geom_smooth(se = FALSE)
ggplot(data = colour_bad, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap( ~ groupvar) ## faceting better
```