---
title: "Section 11: Other Topics"
format: 
  html:
    embed-resources: true
---

## Section 11.2: Large Data

```{r}
library(tidyverse)
diamonds |> nrow()

ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_point()

ggplot(data = diamonds, aes(x = carat, y = price)) +
  geom_hex() +
  scale_fill_viridis_c() +
  theme_minimal()
## can now see that most of the diamonds in this data set
## are low carat and low price (denser region of the hex plot).
```

Really large n (10,000,000 rows)

```{r}
library(tidyverse)
n <- 10000000
x <- rnorm(n, 0, 1)
y <- x * rgamma(n, 3, 2)
df <- tibble(x = x, y = y)
df

##ggplot(data = df, aes(x = x, y = y)) +
##  geom_point()
## above will cause R to crash

## slice_sample randomly chooses 10000 rows to keep
df_small <- df |> slice_sample(n = 10000)
ggplot(data = df_small, aes(x = x, y = y)) +
  geom_point(alpha = 0.2) +
  geom_smooth()
## visualizing a small subset of a very large data set.

## potential downside: missing extreme outliers or data entry errors.
```

Large p:

```{r}
## install.packages("openintro")
library(openintro)
mlb_small <- mlb_players_18 |> filter(games > 50 & position != "P")
mlb_small

library(GGally)
ggpairs(data = mlb_small, columns = 6:11)
## limited to like maybe 6 variables.

## more predictors
ggcorr(data = mlb_small |> dplyr::select(4:ncol(mlb_small)))
## cannot see outliers
## another disadvantage: see Your Turns 1-3. 
```

## Section 11.3: Stat Learning (Skip)

## Section 11.4: Network Visualization

Exercise 1a. Nodes are majors. Edges will correspond to the number of double majors (thicker if there are a lot of double majors and less thick if there are only a few double majors).

Exercise 1b.

a. You are interested in examining how the number of students double majoring at SLU has changed through time.

no: line plot of number of double majors vs. time.

b. You are interested in exploring what other fields statistics majors at SLU major or minor in most often.

network graph could work, but a lollipop plot or bar plot would be more appropriate.

c. You obtain data on all seniors at SLU with an instagram account and want to explore the instagram relationships between different members of the class.

yes.

d. You obtain data on all seniors at SLU with an instagram account and want to explore if there is an association between number of instagram “followers” and GPA.

no: just a scatterplot would work great!


```{r}
library(tidyverse)
library(readxl)
library(here)
df <- read_excel(here::here("data/slu_graduates_17_25.xlsx"))

## fixes error in the data
df <- df |> mutate(across(everything(),
                          .fns = ~replace(., . ==  "STATS" , "STAT")))

majors_long <- df |> filter(!is.na(major3) | !is.na(major2)) |>
  pivot_longer(c(major1, major2, major3),
               names_to = "major_num",
               values_to = "major") |>
  filter(!is.na(major)) |>
  dplyr::select(-minor1, -sex, -minor2, -minor3, -graduation_year)

majors_long <- df |> filter(!is.na(major3) | !is.na(major2)) |>
  pivot_longer(c(major1, major2, major3),
               names_to = "major_num",
               values_to = "major") |>
  filter(!is.na(major)) |>
  dplyr::select(-minor1, -sex, -minor2, -minor3, -graduation_year)

majors_long

network_df <- majors_long |>
  group_by(adm_id) |>
  reframe(test = t(combn(major, 2))) |>
  mutate(vertex1 = test[ ,1],
         vertex2 = test[ ,2]) |>
  dplyr::select(-test)
network_df
```

Exercise 3. If there are triple majors (e.g., HIST, STAT, MATH), we need a row for __each__ combination (HIST, STAT), (HIST, MATH) and (STAT, MATH) all on separate rows.

```{r}

network_sum <- network_df |> mutate(major_first = if_else(vertex1 > vertex2,
                                           true = vertex2,
                                           false = vertex1),
                     major_second = if_else(vertex1 > vertex2,
                                            true = vertex1,
                                            false = vertex2)) |>
  dplyr::select(-vertex1, -vertex2) |>
  group_by(major_first, major_second) |>
  summarise(n_majors = n()) |>
  filter(n_majors >= 7)
```

Exercise 4. it's putting the majors in alphabetical order (e.g., always MATH, PHYS and never PHYS, MATH) so that they get counted up correctly in the summarise().

Exercise 5. filter() ensures that there are not edges between majors that have had fewer than 5 double majors (which will make plot look cleaner).

tidygraph wants two data sets:

a data set that only has the unique nodes, created below, with unique id numbers for each node.
equal to MCSS for MATH, CS, STAT, DATA, and EC_MA,
equal to Science for BIOL, PSYC, and PHYS,
equal to Social Science for ECON, EN_EC, EC_SP, SOC, ANTH, GOVT, and BUSLA
equal to Arts and Humanities for all other majors (can use the .default argument to case_when() for this).
```{r}
nodes <- df |> filter(!is.na(major3) | !is.na(major2)) |>
  pivot_longer(c(major1, major2, major3),
               names_to = "major_num",
               values_to = "major") |>
  filter(!is.na(major)) |>
  distinct(major) |> ## keep only one row for each unique major
  rename(label = major) |>
  ## only keep the majors that are found in network_sum
  filter(label %in% c(network_sum |> pull(major_first),
                      network_sum |> pull(major_second))) |>
  rowid_to_column("id") |>
  mutate(major_type = case_when(label %in% c("MATH", "STAT", "CS", "DATA",
                                             "EC_MA") ~ "MCSS",
                                label %in% c("BIOL", "PSYC", "PHYS") ~ "Science",
                                label %in% c("ECON", "EN_EC", "EC_SP", "SOC", "ANTH", "GOVT", "BUSLA") ~ "Social Science",
                                .default = "Arts and Humanities"))
nodes
```

edges is a data set that has the ids for two vertices along with the variable that controls how thick the edges will be.

```{r}
edges <- left_join(network_sum, nodes, by = join_by(major_first == label)) |>
  rename(from = "id") |>
  left_join(nodes, by = join_by(major_second == label)) |>
  rename(to = "id") |>
  ungroup() |>
  dplyr::select(to, from, n_majors)
edges
```

```{r}
network_obj <- tbl_graph(nodes = nodes, edges = edges, directed = FALSE)
library(ggraph)

ggraph(network_obj |> mutate(centrality = centrality_degree()),
                            layout = "kk", maxiter = 500) +
  geom_edge_link(aes(width = n_majors), alpha = 0.7, show.legend = FALSE) +
  geom_node_point(aes(colour = major_type,
                      size = centrality), show.legend = FALSE) +
  scale_edge_width(range = c(0.5, 3)) +
  geom_node_text(aes(label = label), repel = TRUE) +
  scale_colour_viridis_d()
```

Exercises 8 through 9.

centrality_degree sizes the nodes just according to how many edges each node has.






Your Turn 1-3.

```{r}
set.seed(1414111)
x1 <- rnorm(100, 0, 1)
x2 <- seq(-1, 1, length.out = 100)
y <- 4 + 0 * x1 + 1 * x2^2 + rnorm(100, 0, 0.1)
df <- tibble(x1 = x1, x2 = x2, y = y)
ggcorr(data = df)

## it looks like there isn't a strong relationship between
## x1 and y or between x2 and y
```

```{r}
ggplot(data = df, aes(x = x1, y = y)) +
  geom_point()
ggplot(data = df, aes(x = x2, y = y)) +
  geom_point() 
## it looks like x2 and y actually have a very strong
## relationship.

## your turn 3
## the ggcorr plot will only detect linear associations-
## this is the major disadvantage of using this plot over
## ggpairs()
```

Exercise 4.


```{r}
library(gapminder)
gapminder_df <- gapminder |> filter(year == 2002 | year == 2007)
gapminder_df <- gapminder_df|> mutate(year = factor(year))

## a
ggplot(data = gapminder_df, aes(x = year, y = lifeExp)) +
  geom_boxplot()
## this is flawed because the graph is ignoring how
## the countries are paired (we have dependent data and
## we need to show that dependence!)

## b.
gapminder_wide <- gapminder_df |> pivot_wider(id_cols = country, 
                            names_from = "year",
                            values_from = "lifeExp") |>
  mutate(diff = `2007` - `2002`)

ggplot(gapminder_wide, aes(x = diff)) +
  geom_histogram(colour = "skyblue4", fill = "skyblue1",
                 bins = 15)
## most countries had a positive change between 2002 and
## 2007, but I've lost the actual scale of life expectancies

## c
gapminder_df <- gapminder_df |> group_by(country) |>
  mutate(diff = diff(lifeExp)) |>
  ungroup() |>
  mutate(change = if_else(diff > 0, 
                          true = "inc",
                          false = "dec"))

ggplot(data = gapminder_df, aes(x = year, y = lifeExp)) +
  geom_line(aes(group = country, colour = change), alpha = 0.3) +
  theme_minimal() +
  scale_colour_manual(values = c("#DC3220",
                                 "#005AB5"))
## good for relatively small samples, but if we had 10000 pairs,
## this would be a mess (and would need to do the histogram of differences)
```




