---
title: "Section 12: Ethics"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
theme_set(theme_minimal())
challenger_df <- read_csv(here("data/Challenger.csv"))
```

Exercise 2.

```{r}
library(broom)
launch_mod <- glm(SuccessfulLaunch ~ Temperature, data = challenger_df,
    family = "binomial")
launch_mod |> tidy()
```

According to our model, for a one degree farenheit increase in temperature, the odds of a "successful launch" increase multiplicatively by 1.26 times.

Exercise 3.

```{r}
ggplot(data = challenger_df, aes(x = Temperature,
                                 y = SuccessfulLaunch)) +
  geom_smooth(method = "glm",
              method.args = list(family = "binomial")) +
  geom_rug(data = challenger_df |> filter(SuccessfulLaunch == 1), sides = "t") + 
   geom_rug(data = challenger_df |> filter(SuccessfulLaunch == 0), sides = "b") +
  geom_vline(aes(xintercept = 31), linetype = "dashed")
```

Your Turn 1. Yes, this is a violation of ethics. Even if it was an accident, a couple of principles of data science ethics were violated (similar to if a medical professional made a egregious and negligent error in treating someone, they are still held liable).


