---
title: "Ethics"
author: "Matt Higham"
format: 
  html:
    embed-resources: true
---

```{r}
library(tidyverse)
library(here)
theme_set(theme_minimal())
challenger_df <- read_csv(here("data/Challenger.csv"))
challenger_df
```

Exercise 2.

```{r}
library(broom)
challenger_mod <- glm(SuccessfulLaunch ~ Temperature,
                      data = challenger_df,
                      family = "binomial")
challenger_mod |> tidy()

ggplot(data = challenger_df, aes(x = Temperature,
                                 y = SuccessfulLaunch)) +
   geom_smooth(method = "glm",
               method.args = list(family = "binomial")) +
  geom_rug(data = challenger_df |> filter(SuccessfulLaunch == 1), sides = "t") +
  geom_rug(data = challenger_df |> filter(SuccessfulLaunch == 0), sides = "b") +
  geom_vline(xintercept = 31)
```

Exercise 5. As we can see from the plot, o-rings are less likely to have damage at higher temperatures. Additionally, there were no test flights done at or near 31 degrees farenheit so it would be inadvisable to launch the shuttle at this temperature without further testing.

We agreed that this was an ethical violation, and discussed a violation of principles 8, 9, and 1. Removing the data points led to a poorly informed decision, and more careful consideration of the consequences of something going wrong could have prevented the tragedy.
